\appendix
\chapter{User guide}
\label{cha:user_guide}


\section{List of scripts}

\begin{description}
  \item [create\_jars.sh]
      - it is a script responsible for building jar files of all projects needed to run co-evolutionary algorithm on multi-agent running platform. 
  \item [deploy.sh]
      - script responsible for running co-evolutionary algorithm on a multi-agent platform.
      A special directory called \emph{sandbox} is created and all necessary jar files as well as files containing preprocessed historical data are copied to \emph{sandbox}.
      It also specifies the number of Computing Nodes that will be used for computation.
      ActiveMQ should be started before running this script.
  \item [create\_agent\_jars.sh]
      - script responsible for building jar files of all projects needed to run CoEMAS system on multi-agent running platform. 
  \item [agent\_node\_deploy.sh]
      - script responsible for running CoEMAS system on a multi-agent platform.
      A special directory called \emph{sandbox} is created and all necessary jar files as well as files containing preprocessed historical data are copied to \emph{sandbox}.
      It also specifies the number of Computing Nodes that will be used for computation.
      ActiveMQ should be started before running this script. 
  \item [precompute.R]
      - script responsible for precomputing standard deviation and correlation coefficients from historical data.
  \item [PrecomputeCovVarForMPT.R]
      - script responsible for precomputing covariance and variance from historical data. 

\end{description}



\section{Adjusting algorithms}

\begin{description}
 \item [trend following]
      - in order to adjust trend following algorithm we have to modify the script code.
      Thanks to comments present in code, it is not that difficult.
  \item [genetic algorrithm]
      - there are constants (regulating algorithm behaviour) in file GeneticAlgorithmImpl.java that can be set by class constructor. 
  \item [co-evolutionary algorithm]
      - because Spring manages the bean creation of co-evolutionary algorithm, all coefficients responsible for adjusting algorithm behaviour can be set by constructor injection. 
  \item [CoEMAS] 
      - CoEMAS can be adjusted in the same way as the co-evolutionary algorithm (constructor injection).
\end{description}


\section{Interpreting results}

Running trend following script will produce results at once - graph, showing portfolio value in time, is plotted as soon as computation is finished.
CoEMAS as well as co-evolutionary algorithm are run in a sandbox and results of the computation are stored in a file.
It contains the information about portfolio composition, total value and risk of portfolio, etc. 
Results from genetic algorithm are stored in a similar way. 

