\chapter{Basic Definitions}
\label{cha:basic}


%---------------------------------------------------------------------------

\begin{description}
  \item [JMS]
	Java Message Service, it allows application to create, read, send and receive messages.
	Implementations of JMS provide reliable and asynchronous communication for distributed systems. 

  \item [portfolio]
	Portfolio is a collection of investments owned by an investor.
	It could contain many different assets like stocks or bonds.
	Usually the more diversified the portfolio the safer our investments.

  \item [Inversion of Control]
	Inversion of Control (IoC) is an architectural pattern.
	It relies on concept of having an outside entity (usually it is called IoC container) that is responsible for wiring objects together. \cite{Spring} 

  \item [Spring Framework] 
	Spring Framework is a very popular, open source application development framework that makes the process of   application development much easier.
	It contains many useful modules (e.g. IoC container, transaction management, etc.).
	One of its main advantages is lightweight and flexibility. \cite{Spring}

  \item [Java]
	Object oriented programming language that is portable, secure and robust.
	Java code is compiled to byte code interpeted by Java Virtual Machine. 
	It is very popular thanks to ease of application development. 

  \item [population (statistics)]
	Population is a set containing all entities that are relevant in a statistical study of a particular problem \cite{Oxford}. 

  \item [population mean]
	
	The mean of an observation variable is a numerical measure of the central location of the data values.
	The population mean is defined as:
	 \begin{equation}
			\mu = \frac{1}{N} \sum_{i=1}^{N}{(x_{i})} 
		      \end{equation}
  \item [population variance]
	The variance is a numerical measure of how the data values is dispersed around the mean.
	The population variance is defined as:  
	      \begin{equation}
			\sigma^2 = \frac{1}{N} \sum_{i=1}^{N}{(x_{i} - \mu)} 
		      \end{equation}
	where $\mu$ is a population mean.
	   
  \item [population standard deviation]
	The population standard deviation ($\sigma$) is a square root of the population variance.
  \item [covariance]
	The covariance of two variables x and y in a data sample measures how the two are linearly related.
	A positive covariance would indicates a positive linear relationship between the variables, and a negative covariance would indicate the opposite. 
	The population covariance is defined as:

	\begin{equation}
			\sigma_{xy} = \frac{1}{N} \sum_{i=1}^{N}{(x_{i} - \mu_{x})(y_{i} - \mu_{y})} 
		      \end{equation}
	where $\mu_{x}$ and $\mu_{y}$ are population means.

  \item [correlation coefficient]
	Correlation coefficient is a normalized measurement of how the two variables in a data sample are linearly related.
	The population correlation coefficient is defined as:

	\begin{equation}
	  \rho_{xy} = \frac{\sigma_{xy}}{\sigma_{x} \sigma_{y}}
	\end{equation}
	where $\sigma_{x}$ and $\sigma_{x}$ are the population standard deviations and $\sigma_{xy}$ is the population covariance.
	
  
\end{description}
















